Quantifying the importance of a tree prior in Bayesian phylogenetics

### 2019-04-15 article.tex
Spellcheck

Even when a diversification model passes the procedure described above,
it is not necessarily used in Bayesian inference.
In Bayesian inference, the likelihood of a diversification 
model, is computed millions of times.
Due to this, effective but computationally expensive tree priors are non-viable.
Also, there may be no need to add such a tree prior to a Bayesian
inference framework: the more informative the data, 
the lower the influence the tree prior has.
Due to this, computionally faster tree priors are more commonly used.
However, it is unknown if it is valid to assume that tree prior choice is 
of low importance.

@@

\begin{sidewaystable}
\centering
  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} 
      & \textbf{Possible values} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
      & Yule [\cite{yule}], BD, CBS, CCP, CEP \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates 
      & strict, relaxed log-normal \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{s_{m}}$ & site model & Nucleotide substitution model 
      & JC, HKY, TN93, GTR \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs 
      & $m_{r} \in \mathbb{R}_{>0}$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree 
      & any combination of a, c, g, t \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common 
      ancestor & see documentation \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
      & see documentation \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{t}}$ & model type & Criterion to select an inference model 
      & Generative, Candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{r_{i}}$ & run if & Condition under which an inference model is 
      used & Always, Best candidate\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{e_{f}}$ & error function & Specifies how to measure the error 
      & nLTT, $|\gamma|$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{\tau_{p}}$ & twin tree prior 
      & Sets the tree prior for the twin tree 
      & Yule, BD \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Most important parameter options.
    BD = birth death (\cite{nee1994reconstructed}), 
    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
    CCP = coalescent constant-population, 
    CEP = coalescent exponential-population,
    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
    TN93 = Tamura and Nei \cite{tamura1993estimation}, 
    GTR = Generalised time-reversible model (\cite{tavare1986some}).
  }
  \label{tab:options}
\bigskip

@@

\begin{lstlisting}[
  language=R, floatplacement=ht, frame=single,
  label = {lts:create_pir_params_defaults},
  caption = Create a \texttt{pir\_params} with many defaults.
]
pir_params <- create_pir_params(
  alignment_params = alignment_params,
  experiments = experiments
)
\end{lstlisting}

@@

When you are moderately sure about hand-picking an inference model,
but would also like to check how well your hand-picked model compares
to others, we can do both. Of course, you are interested in the
baseline error for both the hand-picked and best candidate model,
so you'll enable twinning as well. This example is a combination
of the previos two ERQs.

@@

\verb;pirouette;'s is extensively documentated on its website,
its documentation and its vignettes.
The \verb;pirouette; website is a good starting point to learn
how to use \verb;pirouette;, as it links to tutorials and videos.
The \verb;pirouette; package documentation describes
all functions and liberally links to related functions.
All exported functions show a minimal example as part of their documentation.
The \verb;pirouette; vignette demonstrates extensively how 
to use \verb;pirouette; in a more lightly written way.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgements}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%

>>>>

Even when a diversification model passes the procedure described above,
it is not necessarily used in Bayesian inference.
In Bayesian inference, the likelihood of a diversification 
model, is computed millions of times.
Due to this, effective but computationally expensive tree priors are non-viable.
Also, there may be no need to add such a tree prior to a Bayesian
inference framework: the more informative the data, 
the lower the influence the tree prior has.
Due to this, computationally faster tree priors are more commonly used.
However, it is unknown if it is valid to assume that tree prior choice is 
of low importance.

@@

\begin{sidewaystable}
\centering
  \begin{tabular}{|@{}c|p{2.5cm}|p{9cm}|p{4.5cm}@{}|}
    \hline
    \centering
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \textbf{Symbol} & \textbf{Sub-argument} & \textbf{Description} 
      & \textbf{Possible values} \\ 
    \hline
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{t_{p}}$ & tree prior & Macroevolutionary speciation model 
      & Yule [\cite{yule}], BD, CBS, CCP, CEP \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{c_{m}}$ & clock model & Clock for the DNA mutation rates 
      & strict, relaxed log-normal \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{s_{m}}$ & site model & Nucleotide substitution model 
      & JC, HKY, TN93, GTR \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{r}}$ & mutation rate & Pace at which mutation occurs 
      & $m_{r} \in \mathbb{R}_{>0}$\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{r_{s}}$ & root sequence & DNA sequence at the root of the tree 
      & any combination of a, c, g, t \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{p}}$ & MRCA prior & Prior knowledge of a most recent common 
      ancestor & see documentation \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{s}}$ & MCMC settings & Markov Chain Monte Carlo algorithm setup 
      & see documentation \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{m_{t}}$ & model type & Criterion to select an inference model 
      & Generative, Candidate \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{r_{i}}$ & run if & Condition under which an inference model is 
      used & Always, Best candidate\\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{e_{f}}$ & error function & Specifies how to measure the error 
      & nLTT, $|\gamma|$ \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    $\mathit{\tau_{p}}$ & twin tree prior 
      & Sets the tree prior for the twin tree 
      & Yule, BD \\
    %%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
    \hline
  \end{tabular}
  \caption{
    Most important parameter options.
    BD = birth death (\cite{nee1994reconstructed}), 
    CBS = coalescent Bayesian skyline (\cite{drummond2005bayesian}), 
    CCP = coalescent constant-population, 
    CEP = coalescent exponential-population,
    JC = Jukes and Cantor (\cite{jukes1969evolution}), 
    HKY = Hasegawa, Kishino and Yano (\cite{hasegawa1985dating}), 
    TN93 = Tamura and Nei \cite{tamura1993estimation}, 
    GTR = Generalized time-reversible model (\cite{tavare1986some}).
  }
  \label{tab:options}
\bigskip

@@

\begin{lstlisting}[
  language=R, floatplacement=ht, frame=single,
  label = {lst:create_pir_params_defaults},
  caption = Create a \texttt{pir\_params} with many defaults.
]
pir_params <- create_pir_params(
  alignment_params = alignment_params,
  experiments = experiments
)
\end{lstlisting}

@@

When you are moderately sure about hand-picking an inference model,
but would also like to check how well your hand-picked model compares
to others, we can do both. Of course, you are interested in the
baseline error for both the hand-picked and best candidate model,
so you'll enable twinning as well. This example is a combination
of the previous two ERQs.

@@

\verb;pirouette;'s is extensively documented on its website,
its documentation and its vignettes.
The \verb;pirouette; website is a good starting point to learn
how to use \verb;pirouette;, as it links to tutorials and videos.
The \verb;pirouette; package documentation describes
all functions and liberally links to related functions.
All exported functions show a minimal example as part of their documentation.
The \verb;pirouette; vignette demonstrates extensively how 
to use \verb;pirouette; in a more lightly written way.

@@

%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
\section{Acknowledgments}
%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%%
